---
title: "Assignment_3"
author: "Jyothsna P - 811251679"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
###install.packages("e1071")
###loading the required packages:dplyr
###loading the required packages:ISLR
###loading the required packages:tidyverse

library(e1071)
library(dplyr)
library(caret)
library(readr)
library(ISLR)
library(tidyverse)
```

### Read the UniversalBankData to the R environment

```{r}

UniversalBankData_1 <- read.csv("C:/Users/peddi/OneDrive/Desktop/Spring 2023/FML/Module 4/Assignment 2/UniversalBank.csv")

```

### Check the descriptive statistics of universalBank Data

```{r}

summary(UniversalBankData_1)

```

### Partitioning the data to train data 60% and validation data 40%.

```{r}


set.seed(143)

#CreateDataPartition function helps in creating an index to partition data into desired ratio
Index_Train <- createDataPartition(UniversalBankData_1$Personal.Loan,p=0.6,list = FALSE)

#Index_train holds the row index of 60% of data.
TrainData <- UniversalBankData_1[Index_Train,]
ValidationDAta <- UniversalBankData_1[-Index_Train,]

```

# Question: A

```{r}

## Creating pivot table for Online as a column variable, creditcard as row variable and personal loan as a secondary row variab using the table function.

table(CreditCard= TrainData$CreditCard, Online= TrainData$Online, Loan= TrainData$Personal.Loan)


```

# Question: B

### Probability of Loan acceptance of a customer who is owing a bank credit card and actively using online services by using the above pivot table.

```{r}

## Calculating the Probability of accepting loan = Number of observations where Loan = 1 / Total number of observations where CreditCard = 1 and Online = 1.

LoanAcceptProb = 55/(55+492)
LoanAcceptProb

```

# Question: C

## Creating two separate pivot tables for the training data.

### Creating a pivot table for Loan(in rows) and Online(in column) for the training data using the table function.

```{r}

PTforLoanVSOnline = table(Loan= TrainData$Personal.Loan, Online=TrainData$Online)
print(PTforLoanVSOnline)

```

### Creating a pivot table for loan(in rows) as a function of Credit card (as column)for the training data using table function.

```{r}

PTforCCVSLoan = table( Loan= TrainData$Personal.Loan,CC= TrainData$CreditCard)
print(PTforCCVSLoan)
```

# Question: D

```{r}

##Calculating the probability of credit card given by loan when ( CC=1 | Loan=1 )

DI= PTforCCVSLoan[2,2]/
(PTforCCVSLoan[2,1]+PTforCCVSLoan[2,2])

DI

```

```{r}
##Calculating the probability of Online given by Loan when ( Online=1| Loan=1 )

DII= PTforLoanVSOnline[2,2]/
  (PTforLoanVSOnline[2,1]+PTforLoanVSOnline[2,2])

DII

```

```{r}
##Calculating the probability of Loan=1 

# Numerator - Sum of all values where loan=1 - sum(PTforCCVSLoan[2,])
# Denominator - Sum of all values in the matrix - sum(PTforCCVSLoan)

DIII= sum(PTforCCVSLoan[2,])/ sum(PTforCCVSLoan)


DIII

```

```{r}

##Calculating the probability of credit card given by Loan when ( CC=1 | Loan=0 )

DIV = PTforCCVSLoan[1,2]/
  (PTforCCVSLoan[1,2]+PTforCCVSLoan[1,1])
  
DIV
                         

```

```{r}

##Calculating the probability of Online given by Loan when (Online=1| Loan= 0 )

DV = PTforLoanVSOnline[1,2]/
  (PTforLoanVSOnline[1,2]+PTforLoanVSOnline[1,1])

DV

```

```{r}
#Calculating the Probability of Loan= 0 

# Numerator - Sum of all values where loan=0 - sum(PTforCCVSLoan[1,])
# Denominator - Sum of all values in the matrix - sum(PTforCCVSLoan)

DVI = sum(PTforCCVSLoan[1,])/
         sum(PTforCCVSLoan)


DVI


```

# Question: E

## Computing the naive Bayes probability using the quantities computed above for P(Loan = 1 \| CC= 1, Online = 1)

```{r}

E1= (DI*DII*DIII)/((DI*DII*DIII)+(DIV*DV*DVI))

E1


```

# Question: F

## Comparing the probability values obtained in B and in naive Bayes model in E.

### The value obtained in the B is 0.1005484 and value obtained by computing the naive Bayes probability in the E is 0.1011591. By comparing them it is clear that probability obtained from solving naive Bayes model have higher probability than that in B.

# Question: G

```{r}

## Creating the Test data by applying the filters on train data on condition when CC=1 and Online=1

TestData = filter(TrainData, CreditCard== 1 & Online== 1)

head(TestData)

```

### Performing the Naive Bayes model on the training data.

```{r}

nb_Model = naiveBayes(Personal.Loan~ Online + CreditCard, data = TrainData)
nb_Model

```

### Making the predictions and returning the probability values by using the naive Bayes model.

```{r}

Predicted_Test_Model<- predict(nb_Model,TestData, type = "raw")

head(Predicted_Test_Model,n = 8)##showing the first few values from the predicted_test_model


```

### Comparing the probabilities obtained by the Naive Bayes model on training data in E is 0.1011591 and on test data in G is 0.1026888, which are almost same.
