---
title: "FML Assignment 1"
author: "Jyothsna P - 811251679"
date: "1/28/2023"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

# Test Push In GithubDesktop

## R Markdown

##Install Readr Package
#Install readr package to use read.csv file from the local machine 
#repos URL is required to use the install packages in the R markdown file
```{r}
#install.packages("readr",repos = "http://cran.us.r-project.org")
#install.packages("tidyverse")
#library(tidyverse)
#library(readr)
#install.packages("pscl", repos = "https://cran.rstudio.com")

```

##Load Cereals Data
#read.csv function is useful to load the data available in the csv file to r environment
#Cereals.csv file is downloaded from the FML course in Canvas
```{r}
#install.packages("ggplot2")

cerealsData <- read.csv("C:/Users/peddi/Downloads/Cereals.csv",header = TRUE,sep = ",")
```

#Class of the Name column

```{r}

class(cerealsData$name)

```

##Descriptive Statistics

#Check if any null values are there in the cereals dataset
```{r}
any(is.na(cerealsData))
```

#Get Descriptive Statistics of all the columns in the dataset

```{r}
summary(cerealsData)
```

Get Descriptive Statistics of the Calories column in the dataset
```{r}

summary(cerealsData$calories)

```

Transformations

Transform column name type in Cereals dataset to factor

```{r}
cerealsData$type <- factor(cerealsData$type) 
```

```{r}

plot(cerealsData$fiber)
Logtrans_fiber=log10(cerealsData$fiber)
plot(Logtrans_fiber)

```

```{r}
Logtrans_sodium=sqrt(cerealsData$sodium)
Logtrans_sodium
```

Including Plots

Scatter Plot for Calories and fat attributes
```{r}

plot(cerealsData$calories,cerealsData$fat)

```
Plotting ggplot for categorical attribute type and carbo variables

```{r}
library(ggplot2)
ggplot(cerealsData, aes(x=cerealsData$carbo, y=cerealsData$type)) + 
  geom_bar(stat = "identity", width=0.2) 
```

Plotting ggplot for sugars and protein variables

```{r}
library(ggplot2)
ggplot(cerealsData, aes(x=cerealsData$sugars, y=cerealsData$protein)) +
  geom_line()

```
Barplot for sugars variable

```{r}
barplot(cerealsData$sugars)
```

plotting ggplot for mfr and type variables

```{r}
ggplot(cerealsData, aes(mfr, ..count..)) + geom_bar(aes(fill = type), position = "dodge")

```

